### 2020-09-19 manuscript/ERL_global_C_review.Rmd
worked on paper

### Summary
```{r eval=FALSE, echo=FALSE}
# The Summary should be no longer than 300 words and divided into the following sections: Background, Methods/Design, Review results/ Synthesis and Discussion. Up to 7 Keywords should also be provided. The total length of the article is flexible.
```

@@

*Methods/Design.* Here, we draw from the Global Forest C Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation. Specifically, we draw upon ## records from ## geographic locations representing ## C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests. For regrowth forests, we quantify age trends for all variables.

@@

### Background
```{r eval=FALSE, echo=FALSE}
# The Background section needs to present the rationale for why a systematic review of this topic is needed along with a history of what has been done to date and an expectation of what new will emerge from the review, especially if quantitative meta-analyses of studies are being considered.
```

Carbon (C) cycling in forests is central to sustaining life on Earth. Forest ecosystems globally influence climate through their vital role in regulating atmospheric carbon dioxide (CO~2~). Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ (Bagdley et al. 2019), or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**). While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion was retained in ecosystems over recent decades. The resulting in C sink averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**). Moreover, forests contain substantial C stocks: an estimated 92% of terrestrial biomass  (Pan et al 2013) and 45% of terrestrial C (biomass and soils; Bonan 2008). Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), resulting in net C sink of ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**). The future of this sink is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change. Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change [@grassi_key_2017] (Friedlingstein et al 2006, Cavaleri et al 2015, Griscom et al 2017; IPCC1.5).

Despite the centrality of forest C cycling in regulating both atmospheric CO~2~ and ecosystem function and diversity, we lack a comprehensive understanding of how C cycling varies with across biomes and with stand age. While remote sensing measurements are increasingly useful for global- or regional-scale estimates of forest $GPP$ (Bagdley et al. 2019, [@li_mapping_2019]), aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) [@clark_quantifying_2004; @leitold_nino_2018], and to some extent net ecosystem exchange ($NEP$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements. Moreover, data on stand characteristics--including species composition, age, and disturbance history--are critical to interpreting and projecting into the future, but [usually rely on ground-based measurements]. Thus, ground-based measurements are, and will continue to be, central to characterizing forest C cycling.

Tens of thousands of ground-based forest C measurements have been published, but their distribution across literally thousands of scientific articles-- along with variation in data formats, units, measurement methods, *etc.*--have made them effectively inaccessible for many global-scale analyses, including those attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011), and model evaluation (Clark et al 2017, Luo et al 2012). Important progress has been made in synthesizing data to address how C cycling varies across forest biomes (Luyssaert et al. 2007; REFS) and stand ages (REFS), yet generally consider only a limited set of C variables [but see @anderson-teixeira_carbon_2016 for the tropics] and do not consider age-by-biome interactions. To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (@anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018). ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (*e.g.*, stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications. ForC currently contains **#** (~49,000!) records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.

@@

2. How do C cycling vary across the worlds major forest biomes?

3. How does C cycling vary with stand age?

While components of these questions have been previously addressed [@luyssaert_co2_2007; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018, @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.

@@

### Methods/ Design
```{r eval=FALSE, echo=FALSE}
# The Methods/Design section needs to describe how articles in the reviews were identified and what criteria were used for justifying inclusion in the review. While traditional reviews relying on past experience and expert knowledge are acceptable, editors should encourage reviews that are set up as so-called ‘systematic reviews’ (see the Cochrane Review procedure which initiated the process in the medical and health science: http://community.cochrane.org/ about-us/evidence-based-health-care).
```

This review synthesizes data from the ForC database [Fig. 1; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018]. ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies. 
Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records. 
The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age. 
As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible (83% of records in v.2.0; Anderson-Teixeira et al 2018). 
Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], yet these represented only approximately one-third of records in ForC v.2.0 [@anderson-teixeira_forc_2018]. 
Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database (GROA v1.0; Cook-Patton et al. in press; ZenodoDOI), and data from several publications (Taylor et al. 2017, Lutz et al. 2018, Johnson_2018_csss; see GitHub repo for complete list) ([GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv)). 
We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS). 
The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: TBD).

@@

### Review Results/ Synthesis
```{r eval=FALSE, echo=FALSE}
# The Review Results/Synthesis section presents the findings of the analyses.
```
*Data Coverage*

@@

*C cycling in mature forests*

@@

*C cycling in young forests*

@@

### Discussion 
```{r eval=FALSE, echo=FALSE}
# The Discussion section highlights the general conclusions, emphasizes novel findings, puts them in context and identifies issues and inferences for future work or policy considerations.
```

@@

*C variable coverage and budget closure*

@@

*C cycling across biomes*

THere is a general acceleration of carbon cycling from the tropics to the high latitudes. 
For mature forests, this is consistent with lots of previous work, most recently @banbury_morgan_global_nodate.
For regrowth forests, we see faster C accumulation in the tropics (*I assume!- analysis not yet complete*), consistent with the observation that biomass accumulation decreases with latitude (Anderson 2006, Cook-Patton et al. 2020).
For most C flux variables, this analysis is the first to examine flux trends in regrowth forests across biomes  (i.e., age x biome interaction).
Data remain sparse, but for better-represented variables, we often see faster acceleration of C cycling in the warmer climates.
Further work will be required to explore age x climate interactions, but our broad-brush overview indicates that C cycling of regrowth forests is not only higher in the tropics, parallel to fluxes in mature forests [ @banbury_morgan_global_nodate], but also that it accelerates more rapidly with stand age in the tropics, consistent with more rapid biomass accumulation.

*Age trends in C cycling*  (Just some rough notes at this point)

@@

Moreover, as disturbances increase, understanding the carbon dynamics of regrowth forests will be increasingly important.

@@

*Relevance for climate change prediction and mitigation*

@@

Second, ForC can serve as a pipeline through which forest science can inform forest-based climate change mitigation efforts.
Such efforts will be most effective when informed by the best available data, yet it is not feasible for the individuals and organizations designing such efforts to sort through literature, often behind paywalls, with data reported in varying units, terminology, etc. 
One goal for ForC is to serve as a pipeline through which information can flow efficiently from forest researchers to decision-makers working to implement forest conservation strategies at global, national, or landscape scales.
This is already happening! 
ForC has already contributed to updating the IPCC guidelines for carbon accounting in forests (IPCC 2019, Requena-Suarez et al 2019, Rozendaal et al in prep), mapping C accumulation potential from natural forest regrowth globally (Cook-Patton et al. 2020), and informing ecosystem conservation priorities (Goldstein et al. 2020).

There remain numerous data needs for improved accounting of forest carbon stocks and fluxes in GHG accounting.
AGB is the largest stock, and most of the emphasis is on this variable. 
Remote sensing, with calibration based on high-quality ground-based data (Schepaschenko et al. 2019, Chave et al. 2019), is the best approach for mapping forest carbon (REFS).
However, it is limited in that it is not associated with stand age and disturbance history, except in recent decades (e.g., Hansen et al.; REFS).
ForC  is therefore valuable in defining age-based trajectories in biomass, as in Cook-Patton et al. 2020.

@@

We recommend that use of ForC data go to the original database, as opposed to useing "off-the-shelf" values from this publication. 
This is because
(1) ForC is constantly being updated,
(2) analyses should be designed to match the application,
(3) age equations presented here all fit a single functional form that is not necessarily the best possible for all the variables.

@@

### Acknowledgements
All researchers whose data is included in ForC and this analysis.
Thanks to Norbert Kunert and [Helene's intern] for helpful input at an earlier phase.
A Smithsonian Scholarly Studies grant to KAT and HML. 
WLS grant to KAT.

### References

>>>>

## Summary
```{r eval=FALSE, echo=FALSE}
# The Summary should be no longer than 300 words and divided into the following sections: Background, Methods/Design, Review results/ Synthesis and Discussion. Up to 7 Keywords should also be provided. The total length of the article is flexible.
```

@@

*Methods/Design.* Here, we draw from the Global Forest C Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation. Specifically, we draw upon *##* records from *##* geographic locations representing *##* C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests. For regrowth forests, we quantify age trends for all variables.

@@

## Background
```{r eval=FALSE, echo=FALSE}
# The Background section needs to present the rationale for why a systematic review of this topic is needed along with a history of what has been done to date and an expectation of what new will emerge from the review, especially if quantitative meta-analyses of studies are being considered.
```

Forest ecosystems globally play a critical role in regulating atmospheric carbon dioxide (CO~2~), and thereby will shape the course of climate change (IPCC1.5). 
Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**). 
While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion was retained in ecosystems over recent decades. 
The resulting CO~2~ sink averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**). 
Moreover, forests contain substantial carbon (C) stocks: an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C [biomass and soils; @bonan_forests_2008]. 
Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), resulting in net C sink of ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**). 
The future of this sink is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change. 
Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change [@grassi_key_2017; Cavaleri et al 2015, Griscom et al 2017].

Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [Friedlingstein et al 2006; @krause_large_2018; REFS] and CO~2~ accounting frameworks [Pan, REFS] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history. 
These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published. 
While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [@li_mapping_2019; **REFS for biomass, biomass change, net CO2 flux**], measurement of most forest C stocks and fluxes require intensive on-the-ground data collection. 
Original studies typically cover only a small numbers of sites at a time, with rare exceptions spanning regions or continents [e.g., @lutz_global_2018; FLUXNET_REF), typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015] or FLUXNET [@baldocchi_fluxnet_2001].
The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*. 
In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle (*e.g.*, Pan et al 2011), or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere (REFS).

To address the need for global-scale analyses of forest C cycling, we recently developed an open-access Global Forest Carbon database, ForC (@anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018). 
ForC contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables) based on ground-based measurements, along with associated data required for interpretation (*e.g.*, stand history, measurement methods).
These data have been amalgamated from original peer-reviewed publications, either directly or via intermediary data compilations, with notable recent additions of the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020]. 
ForC currently contains **#** (~49,000!) records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.

@@

2. How do C cycling vary across the world's major forest biomes?

3. How does C cycling vary with stand age (in interaction with biome)?

While components of these questions have been previously addressed [@luyssaert_co2_2007; @anderson-teixeira_carbon_2016; @cook-patton_mapping_2020; @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.

@@

## Methods/ Design
```{r eval=FALSE, echo=FALSE}
# The Methods/Design section needs to describe how articles in the reviews were identified and what criteria were used for justifying inclusion in the review. While traditional reviews relying on past experience and expert knowledge are acceptable, editors should encourage reviews that are set up as so-called ‘systematic reviews’ (see the Cochrane Review procedure which initiated the process in the medical and health science: http://community.cochrane.org/ about-us/evidence-based-health-care).
```

This review synthesizes data from the ForC database [Fig. 1; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018]. 
ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies. 
Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records. 
The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age. 
As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible [83% of records in v.2.0; @anderson-teixeira_forc_2018]. 
Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], yet these represented only approximately one-third of records in ForC v.2.0 [@anderson-teixeira_forc_2018]. 
Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database (GROA v1.0; Cook-Patton et al. in press; ZenodoDOI), and data from several publications (Taylor et al. 2017, Lutz et al. 2018, Johnson_2018_csss; see GitHub repo for complete list) ([GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv)). 
We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS). 
The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: TBD).

@@

## Review Results/ Synthesis
```{r eval=FALSE, echo=FALSE}
# The Review Results/Synthesis section presents the findings of the analyses.
```
### Data Coverage

@@

### C cycling in mature forests

@@

### C cycling in young forests

@@

## Discussion 
```{r eval=FALSE, echo=FALSE}
# The Discussion section highlights the general conclusions, emphasizes novel findings, puts them in context and identifies issues and inferences for future work or policy considerations.
```

@@

### C variable coverage and budget closure

@@

### C cycling across biomes

Our analysis reveals a general acceleration of carbon cycling from the tropics to the high latitudes. 
For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude [e.g.,  @banbury_morgan_global_nodate].
For regrowth forests, more rapid accumulation of biomass at lower latitudes has been well-established [@anderson_temperature-dependence_2006; @cook-patton_mapping_2020], whereas this is the first study to compare age trends in deadwood and organic layer across biomes [but see @cook-patton_mapping_2020].
For most C flux variables, this analysis is the first to examine flux trends in regrowth forests across biomes  (i.e., age x biome interaction).
Data remain sparse, but for better-represented variables, we often see faster acceleration of C cycling in the warmer climates.
Further work will be required to explore age x climate interactions, but our broad-brush overview indicates that C cycling of regrowth forests is not only higher in the tropics, parallel to fluxes in mature forests [ @banbury_morgan_global_nodate], but also that it accelerates more rapidly with stand age in the tropics, consistent with more rapid biomass accumulation.

In contrast to C fluxes and accumulation rates in regrowth forests, stocks...

### Age trends in C cycling

*(Just some rough notes at this point)*

@@

Moreover, as disturbances increase [@andela_human-driven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests will be increasingly important.

It's also important to understand secondary forest C sequestration to reduce uncertainty regarding the potential for carbon uptake by regrowth forests [@kruase_large_2018; @cook-patton_mapping_2020].

@@

### Relevance for climate change prediction and mitigation

@@

Second, ForC can serve as a pipeline through which forest science can inform forest-based climate change mitigation efforts.
Such efforts will be most effective when informed by the best available data, yet it is not feasible for the individuals and organizations designing such efforts to sort through literature, often behind paywalls, with data reported in varying units, terminology, etc. 
One goal for ForC is to serve as a pipeline through which information can flow efficiently from forest researchers to decision-makers working to implement forest conservation strategies at global, national, or landscape scales.
This is already happening! 
ForC has already contributed to updating the IPCC guidelines for carbon accounting in forests [IPCC 2019; @requena_suarez_estimating_2019; Rozendaal et al in prep], mapping C accumulation potential from natural forest regrowth globally [@cook-patton_mapping_2020], and informing ecosystem conservation priorities [@goldstein_protecting_2020].

There remain numerous data needs for improved accounting of forest carbon stocks and fluxes in GHG accounting.
AGB is the largest stock, and most of the emphasis is on this variable. 
Remote sensing, with calibration based on high-quality ground-based data [@schepaschenko_forest_2019; @chave_ground_2019], is the best approach for mapping forest carbon (REFS).
However, it is limited in that it is not associated with stand age and disturbance history, except in recent decades when satellite data can be used to detect forest loss, gain, and some of their dominant drivers [@hansen_high-resolution_2013; @song_global_2018; @curtis_classifying_2018].
ForC  is therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020.

 *remote sensing measurements are increasingly useful for global- or regional-scale estimates of forest $GPP$ (Bagdley et al. 2019, [@li_mapping_2019]), aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) [@clark_quantifying_2004; @leitold_nino_2018], and to some extent net ecosystem exchange ($NEP$) (REFS),*

@@

We recommend that use of ForC data go to the original database, as opposed to using "off-the-shelf" values from this publication. 
This is because
(1) ForC is constantly being updated,
(2) analyses should be designed to match the application,
(3) age equations presented here all fit a single functional form that is not necessarily the best possible for all the variables.

@@

## Acknowledgements
All researchers whose data is included in ForC and this analysis.
Thanks to Norbert Kunert and [Helene's intern] for helpful input at an earlier phase.
A Smithsonian Scholarly Studies grant to KAT and HML. 
WLS grant to KAT.

## References
